---
title: "Step 4: Analysing BIO-WELL scores"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 4: Analysing BIO-WELL scores}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# <span style="color:#0097AD;">Introduction</span>

After your BIO-WELL survey URL has been shared to participants and their
responses have been remotely stored to your Dropbox App's folder, it is time to
move to the analysis stage.

In this vignette, we will be using the `BIOWELL` package functions
`download_data()`, `evaluate_survey()` and `explore_data()` to download the
survey response data, evaluate your survey and explore patterns in BIO-WELL
scores.


# <span style="color:#0097AD;">Step 1: Download survey response data</span>

Participant responses have been saved as individual `csv` files within your
designated Dropbox App's folder.

The `BIOWELL` package function `download_data()` collates all survey responses
and returns a single data frame for analysis.

```{r, eval = FALSE}
path_to_folder <- tempdir() # Replace with path to a local folder. 
appName <- "my_app_name" # Replace with name for your app. 
dir.create(paste0(path_to_folder,"/",appName)) 


survey_data <- download_data(Dropbox_App_folder = "survey_responses",
                             BW_app_path = paste0(path_to_folder,"/",appName))

```

# <span style="color:#0097AD;">Step 2: Evaluate survey </span>

When analysing survey response data, it is important to evaluate the reliability
of your survey.  

The function `evaluate_survey()` offers multiple methods to evaluate the
reliability of surveys.

1) **<span style="color:#004A86;">Cronbach's alpha</span>**

The computed Cronbach's Alpha statistic for your survey,
which measures the internal consistency of responses. This calculated using
`cronbach.alpha()` function from the `ltm` R package (Rizopoulos et al.,
2007).

2) **<span style="color:#004A86;">Item-total correlation</span>**

Item-total correlation: the computed item-total statistic for each BIO-WELL
question within your survey (Cronbach et al., 1951). This is calculated using
the `item.total()` function from the `multilevel` [R
package](https://CRAN.R-project.org/package=multilevel).

Cronbach, L. J. (1951) Coefficient alpha and the internal structure of tests.
Psychometrika, 16, 297–334.
```{r}
library(BIOWELL)
data("sample_BW_data")
results <- evaluate_survey(sample_BW_data)
results[[2]]
results[[3]]
```

# <span style="color:#0097AD;">Step 3: Explore survey responses</span>

To facilitate preliminary analyses of survey response data, the `BIOWELL`
package includes the function `explore_data()`.

The arguments of the function are:

+ `data` - The download survey response data frame (from Step 1!).

+ `var_name` - A character string, text for the y-axis of the returned box plot.

+ `plot` - A logical, whether you want the box plots to be displayed in the R
plot window.

+ `column_name` - An optional argument, the name of a column within data to
split BIO-WELL scores by in the returned box plot.

+ `biowell_scale` - An optional argument, the name of one wellbeing domain with the BIO-WELL scale. If specified, plotted BIO-WELL scores only incorporate the slider
input for this domain. One of; `physical`, `emotional`, `cognitive`, `social`
and `spiritual`.


1) **<span style="color:#004A86;">Default</span>**

If you only provide `explore_data()` with the participant survey response data
frame, by default, the function will collate the BIO-WELL scores for each
participant by situation and question. The function returns the processed data
frame and generates a box plot for each situation to display the BIO-WELL scores
for each question within that situation.

```{r fig.width=7, fig.height=10}
results <- explore_data(data = sample_BW_data,
                        plot = FALSE)
results[[1]][1]
```

Using the argument `biowell_scale`, you could alter the default to plot the
BIO-WELL score calculated from only one well-being domain.


```{r fig.width=7, fig.height=10}
results <- explore_data(data = sample_BW_data,
                        biowell_scale = "emotional",
                        plot = FALSE)
results[[1]][1]
```

2) **<span style="color:#004A86;">Column</span>**

The survey response data frame contains multiple columns including for responses
to any start or end questions.

Using the `column_name` argument in `explore_data()` you can specify a column
that contains categorical data to split participant BIO-WELL scores by instead
of by situation/question (the default).

For example, the below code will generate a box plot comparing BIO-WELL
scores between each age category of participants.

```{r fig.width=7}
results <- explore_data(data = sample_BW_data,
                        column_name = "Age",
                        var_name = "Participants age range",
                        plot = TRUE)
results <- explore_data(data = sample_BW_data,
                        column_name = "Urbanicity",
                        var_name = "Whether participants live in Urban or Rural areas",
                        plot = TRUE)
```


Using the argument `biowell_scale`, you could further alter the `column_name`
plot to plot the BIO-WELL score calculated from only one wellbeing domain.


```{r fig.width=7}
results <- explore_data(data = sample_BW_data,
                        column_name = "Age",
                        var_name = "Participants age range",
                        biowell_scale = "cognitive",
                        plot = TRUE)
```

# <span style="color:#0097AD;">Further analyses</span>

We hope the functions covered in this Vignette facilitate preliminary analysis
of BIO-WELL scores from your `BIOWELL` package surveys.

This concludes our vignettes for the `BIOWELL` package. Please visit the
package's [GitHub page](https://github.com/r-a-dobson/BIOWELL) to report any
bugs or to ask questions on how to use the package.


