---
title: "Vignette E: Evaluate and explore BIO-WELL scores"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette E: Evaluate and explore BIO-WELL scores}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# <span style="color:#0097AD;">Introduction</span>

Once you have downloaded your survey response data (Vignette D), you can use the
`BIOWELL` package functions `evaluate_survey()` and `explore_data()` to evaluate
your survey and explore patterns in BIO-WELL scores.


# <span style="color:#0097AD;">Step 1: Load sample survey response data </span>

In this vignette, we will be demonstrating the `evaluate_survey()` and `explore_data()` on the BIOWELL package in-built dataset.

Run the following code to load this dataset into your local environment. 

```{r, eval = FALSE}
library(BIOWELL)

data("sample_BW_data")
```


# <span style="color:#0097AD;">Step 2: Evaluate survey </span>

When analysing survey response data, it is important to evaluate the reliability
of your survey.  

The function `evaluate_survey()` offers multiple methods to evaluate the
reliability of surveys.

1) **<span style="color:#004A86;">Cronbach's alpha</span>**

The computed Cronbach's Alpha statistic measures the reliability, or internal consistency, of the five wellbeing sliders. This can be calculated for each biodiversity statement and will help understand the strength or consistency with which the sliders collectively measure a biopsychosocial-spiritual model of wellbeing. The resulting alpha statistic can be compared to those obtained in the original Irvine et al. (2023) article. This is calculated using `cronbach.alpha()` function from the `ltm` R package (Rizopoulos et al., 2007).

2) **<span style="color:#004A86;">Item-total correlation</span>**

The computed item-total statistic for each BIO-WELL question within your survey (Cronbach et al., 1951). This assesses how strongly scores on one wellbeing slider are related to scores on all the other wellbeing sliders. It also contributes to understanding the reliability of the wellbeing sliders as a measure of a biopsychosocial-spiritual model of wellbeing. This is calculated using the `item.total()` function from the `multilevel` [R package](https://CRAN.R-project.org/package=multilevel).

Cronbach, L. J. (1951) Coefficient alpha and the internal structure of tests.
Psychometrika, 16, 297–334.
```{r}
results <- evaluate_survey(sample_BW_data)
results[[2]]
results[[3]]
```

# <span style="color:#0097AD;">Step 3: Explore survey responses</span>

To facilitate preliminary analyses of survey response data, the `BIOWELL`
package includes the function `explore_data()`.

The arguments of the function are:

+ `data` - The download survey response data frame (from Step 1!).

+ `var_name` - A character string, text for the y-axis of the returned box plot.

+ `plot` - A logical, whether you want the box plots to be displayed in the R
plot window.

+ `column_name` - An optional argument, the name of a column within data to
split BIO-WELL scores by in the returned box plot.

+ `biowell_scale` - An optional argument, the name of one wellbeing domain with the BIO-WELL scale. If specified, plotted BIO-WELL scores only incorporate the slider
input for this domain. One of; `physical`, `emotional`, `cognitive`, `social`
and `spiritual`.

+ `score` - An optional argument, indicating whether to plot the raw BIO-WELL scores or inverted BIO-WELL scores. One of; `inverted` or `raw`. 

By default, the typical, inverted BIO-well score is plotted. This is calculated
by taking the average score from across five wellbeing domains (physical,
emotional, cognitive, social and spiritual). When `inverted`, the BIO-WELL score
can be interpreted as 0-49, a negative wellbeing response to biodiversity, 50, a
neutral wellbeing responses and 51-100 representing a positive wellbeing
response to biodiversity. The `raw` BIO-WELL scores represents the opposite of
this with 0-49, a positive wellbeing response to biodiversity, 50, a neutral
wellbeing responses and 51-100 representing a negative wellbeing response.

1) **<span style="color:#004A86;">Default</span>**

If you only provide `explore_data()` with the participant survey response data
frame, by default, the function will collate the BIO-WELL scores for each
participant by setting and question. The function returns the processed data
frame and generates a box plot for each setting to display the BIO-WELL scores
for each question within that setting.

**Note: By default this function plots the inverted BIO-WELL scores.** 

```{r fig.width=7, fig.height=10}
results <- explore_data(data = sample_BW_data,
                        plot = FALSE)
results[[1]][1]
```

However, the argument `score` enables you to plot the raw BIO-WELL scores if
desired.

```{r fig.width=7, fig.height=10}
results <- explore_data(data = sample_BW_data,
                        plot = FALSE,
                        score = "raw")
results[[1]][1]
```


Using the argument `biowell_scale`, you could alter the default to plot the
BIO-WELL score calculated from only one well-being domain.


```{r fig.width=7, fig.height=10}
results <- explore_data(data = sample_BW_data,
                        biowell_scale = "emotional",
                        plot = FALSE)
results[[1]][1]
```

2) **<span style="color:#004A86;">Column</span>**

The survey response data frame contains multiple columns including for responses
to any start or end questions.

Using the `column_name` argument in `explore_data()` you can specify a column
that contains categorical data to split participant BIO-WELL scores by.

For example, the below code will generate a box plot comparing BIO-WELL
scores between each age category of participants.

```{r fig.width=7}
results <- explore_data(data = sample_BW_data,
                        column_name = "Age",
                        var_name = "Participants age range",
                        plot = TRUE)
results <- explore_data(data = sample_BW_data,
                        column_name = "Urbanicity",
                        var_name = "Whether participants live in Urban or Rural areas",
                        plot = TRUE)
```


Using the argument `biowell_scale`, you could further alter the `column_name`
plot to plot the BIO-WELL score calculated from only one wellbeing domain.


```{r fig.width=7}
results <- explore_data(data = sample_BW_data,
                        column_name = "Age",
                        var_name = "Participants age range",
                        biowell_scale = "cognitive",
                        plot = TRUE)
```

# <span style="color:#0097AD;">Further analyses</span>

We hope the functions covered in this Vignette facilitate preliminary analysis
of BIO-WELL scores from your `BIOWELL` package surveys.

This concludes our vignettes for the `BIOWELL` package. Please visit the
package's [GitHub page](https://github.com/r-a-dobson/BIOWELL) to report any
bugs or to ask questions on how to use the package.


